### 2018-01-20 analysis/paper/paper.Rmd
up to choosing the best underlying model

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(modelextinctionideas)
```

# Introduction

Here is a citation [@Marwick2017]

# Background

```{r load-data}

# #read in the data, resulting from the searchof Web of Science for TOPIC = "Archaeology", document type = "Article", all years

# reload --------------------------------------------------------------------
# this is the fast way to resume:
items_df <- readRDS(paste0(here::here(),"/analysis/data/derived_data/saa2018-data-df.rds"))

library(tidyverse)
items_df <-
items_df %>%
  # remove branding
  filter(authors != "FN Clarivate Analytics Web of Science\nVR 1.0") %>%
  # uniques only
  group_by(authors, title, journal) %>%
  filter(row_number() == 1)

items_wth_refs <-
  items_df %>%
  filter(!is.na(refs)) %>%
  filter(!is.na(year))

# df of refs with year that each ref appears, and ID of citing article
library(stringr)
library(purrrlyr)

# get data frame with refs for each article
# each row is a single citation,
# with info of article that has that citation
# the cited article is in the .out column
# xx <-
# items_wth_refs %>%
#   by_row(..f = function(this_row){
#    unlist(str_split(this_row$refs, "\n"))
#   }) %>%
#   unnest(.out) %>%
#   # remove non-refs
#   filter(!grepl("^[A-Z0-9]{2} .*|^ER$", .out)) %>%
#   filter( .out != "") %>%
#   mutate(.out = trimws(toupper(.out))) %>%
#   mutate(.out = str_replace_all(.out, "[[:punct:]]", ""))
# 
# # extract year of cited work
# xx1 <- 
# xx %>% 
#   mutate(cited_year = as.numeric(str_extract(.out, "[0-9]{4}"))) %>% 
#   mutate(cited_year = ifelse(cited_year %in% 1900:2017, 
#                              cited_year,
#                              NA))

# make sure that a cited item has a year that is earlier than the year of the
# item that cited it. Because sometimes there are in press items with the following
# year, we +1 
# xx2 <- 
#   xx1 %>% 
#   filter(cited_year <= year + 1)

# # # # save this as rds to save time
#  saveRDS(xx2,
#            paste0(here::here(), "/analysis/data/derived_data/items_wth_refs.rds"))

# If we are resuming the analysis we can just start from here, and read in the data
# frame of citations.
xx <- readRDS( paste0(here::here(), "/analysis/data/derived_data/items_wth_refs.rds"))
```

## Basic exploration of the data

```{r}

# how many items do we have?
hm_items <- nrow(items_wth_refs) # 12133

# span what time period?
time_range <- range(items_wth_refs$year) # 1905 2017

```

We have `r hm_items` articles in our study. The publication dates range from `r time_range[1]` to  `r time_range[2]`. The bibliographic data for these articles was obtained from a search of Clarivate Analytics Web of Science, using the keyword 'archaeolog' and filtering the results to keep only research articles.

The distribution of articles is strongly skewed to more recent years, with a rapid increase after the mid-2000s. For the rest of our analysis we're going to limit the time frame to 1975-207 because the data from beofre 1975 is incomplete for some key journals, such as _American Antiquity_.

```{r}
items_wth_refs %>% 
  group_by(year) %>% 
  tally() %>% 
ggplot(aes(year,
           n)) +
  geom_col() +
  theme_bw() +
  ggtitle("Archaeology articles per year") +
  labs(caption = "Data from Clarivate Analytics Web of Science")

items_wth_refs %>% 
  filter(year >= 1975) %>% 
  group_by(year) %>% 
  tally() %>% 
ggplot(aes(year,
           n)) +
  geom_col() +
  theme_bw() +
  ggtitle("Archaeology articles per year (after 1975)") +
  labs(caption = "Data from Clarivate Analytics Web of Science")
```

Let's look at the citations in each of these articles. As expected, the number of citations per year closely follows the distribution of articles per year.

```{r}
# how many citations per year? For only after 1975
xxx <-
xx %>%
  filter(year >= 1975) %>% 
  group_by(year) %>%
  tally()

# # plot, why do we have some gaps? in the 1970s? This is a gap in the Clarivate database
# ggplot(xxx,
#        aes(year,
#             n)) +
#   geom_col() +
#   #scale_y_log10() +
#   theme_bw() +
#   ggtitle("Citations per year in archaeology articles")

library(scales)
ggplot(xxx,
       aes(year,
            n)) +
  geom_col() +
  scale_y_log10(breaks = trans_breaks('log10', function(x) 10^x),
                  labels = trans_format('log10', math_format(10^.x))) +
  theme_bw() +
  ggtitle("Citations per year in archaeology articles (since 1975)")
```

With such a large data set we can answer a few basic interesting questions, for example:

- What is the most cited paper?    
- What is the most cited paper per decade?

```{r}
most_cited <- 
  xx %>% 
  filter(year >= 1975) %>% 
  count(.out, sort = TRUE)  %>% 
  slice(1:10)
```

```{r}
# from https://gist.github.com/benmarwick/5826552
## create new column that is 'decade'
# first make a lookup table to get a decade for each individual year
year1 <- 1900:2050
my_seq <- seq(year1[1], year1[length(year1)], by = 10)
indx <- findInterval(year1, my_seq)
ind <- seq(1, length(my_seq), by = 1)
labl1 <- paste(my_seq[ind], my_seq[ind + 1], sep = "-")[-42]
dat1 <- data_frame(year = year1, 
                   decade = labl1[indx])
# merge the decade column onto my_df
my_df <- left_join(xx, dat1, by = 'year')

most_cited_per_decade <- 
  my_df %>% 
  filter(year >= 1975) %>% 
  group_by(decade) %>% 
  count(.out, sort = TRUE) %>% 
  filter(n == max(n))
```

We can zoom in on one article, Binford's "Archaeology as Anthropology", published in 1962. The plot shows the proportion of all citations in each year that are Binford 1962.

At a glance, we have three phases, roughly, high-frequency citations during 1962-1980, medium frequency during 1980-1990, and low-frequency 1990-present. This pattern suggest that this paper might eventually go 'extinct' in the literature and not longer be cited.

We can model rates of extinction using tools from palaeontology.

```{r}

# age distribution of "Archaeology as Anthropology" citations, as a proportion
# of all citations in each year
binford_1962_cites_per_year <-
  xx %>%
  filter(year >= 1975) %>% 
  filter(grepl("BINFORD LR 1962 AM ANTIQUITY V28 P217", 
               .out)) %>%
    group_by(year) %>%
    summarise(count = n()) %>%
    left_join(xxx, 
              by = "year") %>%
    mutate(prop = count / n) %>%
    distinct(year, 
             prop, 
             n)

# plot, why no citations throughout the 1970s?
binford_1962_cites_per_year %>%
  ggplot(aes(year, prop)) +
  geom_col() +
  theme_bw() +
  ylab("Proportion of all\ncitations per year") +
  ggtitle("Citations per year of Binford 1962 (since 1975)")
```

# Innovation and extinction rate modelling

## Preparing the data

### Setting the time unit

These citation data are structured so that we have multiple dates of occurance for each paper. We call this an 'occurrence data' set (as we have dates for each individual occurrence), which can be contrasted with a 'lineage' data set where we'd have only the dates of the first and last instance of each article.

To compute extinction and innovation rates on these data we must change the time scale so that 0 is the most recent time unit and the higher the number the further back in time.

```{r}
max_year <- max(xx$year)

xx0 <- 
  xx %>% 
  filter(year >= 1975) %>% 
  mutate(time_unit = max_year - year)

```

### Formatting Ocurrence Data for PyRate

Data sets in occurrence format, such as we have here, provide the opportunity to incoporate uncertainty about the preservation of the observations when analyzing origination and extinction rates. If a researcher does not have a high level of certainty that the dates of first and last occurrence are correct, such as is often the case with fossils or artifacts,  preservation can be informative [@silvestro2014].

We need a data frame with four columns:

Species: which refers the smallest taxonomic unit, here this is each cited reference. In an archaeological context this would be something like a pottery sherd.

Status: which identifies whether the reference is still used today, we will consider all to be 'extinct'. This would also be true of archaeological artefacts.

min_age: this refers to the most recent year that a citation appeared, in years before 2017

max_age: this refers to the earliest year that a citation appeared, in years before 2017

trait: this refers to some discrete trait that is known about each entry. In this example, it is the citing article. In an archaeological example, each trait could represent an archaeological site in which the sherd was found at.

```{r, eval=TRUE}

# must be in this order: c("Species", "Status", "min_age", "max_age", "trait")

refs_occurance_format <- 
xx0 %>% 
  mutate(citing_paper = paste0(authors, 
                               " ",
                               year,
                               " ",
                               title,
                               " ",
                               journal)) %>% 
  select(citing_paper,
         .out, 
         time_unit) %>% 
  group_by(.out) %>%
  mutate(min_age = min(time_unit), 
         max_age = max(time_unit),
         Status = "extinct") %>% 
  rename(Species = .out,
         trait = citing_paper
         ) %>% 
    select(Species,
           Status,
           min_age,
           max_age,
           trait)

head(refs_occurance_format)
```

We can make a histogram showing the distribution of the number of years each cited article was cited for (artcle lifespan):

```{r}
refs_occurance_format_hist <- 
  refs_occurance_format %>% 
  mutate(number_of_years_cited = max_age - min_age + 1)

# ggplot(refs_occurance_format_hist,
#          aes(number_of_years_cited)) +
#   geom_histogram()

hist(refs_occurance_format_hist$number_of_years_cited,
     main = paste("Number of years each cited article was cited for"),
     xlab = "Number of years cited")
```

And what about the paper published in each 5-year interval of our data that has been cited over the longest duration? In fact there are many ties here, so we've just randomly sampled from each 5 year interval.

```{r}
# set up lookup table for 5-year intervals
year1 <- 1975:2020
my_seq <- seq(year1[1], year1[length(year1)], by = 5)
indx <- findInterval(year1, my_seq)
ind <- seq(1, length(my_seq), by = 1)
labl1 <- paste(my_seq[ind], my_seq[ind + 1], sep = "-")[-42]
dat5 <- data_frame(year = year1, 
                   five_year = labl1[indx])

refs_occurance_format_hist_year <- 
refs_occurance_format_hist %>% 
  mutate(year = as.numeric(str_extract(Species, "[0-9]{4}"))) %>% 
  mutate(year = if_else(year %in% 1975:2017, year, 0)) %>% 
  filter(year >= 1975) %>% 
   left_join(dat5, by = 'year') %>% 
  group_by(five_year) %>% 
  filter(number_of_years_cited == max(number_of_years_cited)) %>% 
  distinct(Species, 
           number_of_years_cited, 
           year) %>% 
  slice(1)

refs_occurance_format_hist_year %>% 
  mutate(end_year = year + number_of_years_cited) %>% 
  ungroup() %>% 
  select(Species, year, end_year) %>% 
  ggplot() +
  geom_linerange(aes(reorder(x = Species, year),
                     ymin = year,
                     ymax = end_year)) +
    coord_flip() +
  ylim(c(1975, 
         2019))

```

## Bayesian estimation of diversification

From https://www.nature.com/articles/palcomms201619:

The diversification of citations through time is modelled here as the product of both origination and extinction. The estimation of origination and extinction rates utilizes an underlying birth–death model, which assumes these processes are random events occurring through continuous time. The probability of events occurring at any given time is determined by parameters of the birth–death process, which express the expected number of origination (or extinction) events per lineage per units of time. Rates can vary through time by introducing rate shifts (Silvestro et al., 2014b; Stadler, 2011; Stadler, 2012).

The first year that an article is cited is considered as an origination event, with the lastyear it was cited is characterized as an extinction event.

The next step is to prepare this occurance data frame for input into Python to compute estimations of the origination, extinction and preservation rates. This is accomplished by calling the `extract.ages` function which extracts the ages and creates a readable python file:

```{r, eval=TRUE}
library(pyrater)

# can't do it with occurance format: Error: cannot allocate vector of size 29.5 Mb 
# so let's use the lineage format, where each citation only appears once
refs_lineage_format <- 
  refs_occurance_format %>% 
  group_by(Species) %>% 
  summarise(first_year = min(min_age),
            last_year = max(max_age))

refs_lineage_format$clade <- 0

refs_species_pyrate_data <- lineage_pyrate(refs_lineage_format)

refs_species_pyrate_data$species_pyrate_data$clade <- 0

#  We can preview each data frame:
lapply(refs_species_pyrate_data, head)

```

We write a text file to be used with the PyRate analysis later on. This file will appear in our current working directory

```{r, eval=TRUE}
write.table(refs_species_pyrate_data$species_pyrate_data,
            "lineage_pyrate_data.txt",
            quote=FALSE,
            sep="\t",
            row.names = FALSE)
```

A python file titled `occurrence_pyrate_data_PyRate.py` should have been created and available in the working directory.

# Summary Plots

Prior to starting the PyRate analysis, it is good practice to get a sense of the data by developing a series of plots.  Using the function below, the following plots are created

1. Log Lineages Through Time
2. Cumulative Diversity Through Time
3. Average Lifespan Through Time
4. Histogram of Lifespans

### Citation Data (Lineage Data)

```{r, eval=TRUE, fig.height=10, fig.width=14}
# The numeral indicates the x-axis increments with 
# the number 5 indicating 5-year time bins
# windows() # if interactive
summary_plots(refs_species_pyrate_data$species_pyrate_data, 5) 
```

# Methods

# Results

# Discussion

# Conclusion

# Acknowledgements

# References

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```