### 2018-03-17 analysis/paper/citation-clusters/How-to-cluster-citations.Rmd
compute on citation clusters to get initial pyrate output

## Clustering citations into meaningful groups

For the citations to be useful in the modelling, we need to put them into groups of topics. This is challenging because the citations are so fragmented - we don't have the complete text of the titles in the citations so we can't use typical text analysis methods like clustering by words or topic modelling.

How many unique citations do we have?

```{r}
length(unique(xx$.out))
```

How can we put them into groups?

One approach we can try is to identify clusters of similar citations according to their co-occurances in citation lists of articles. The idea is that similar articles will cite similar items. So if we can identify similar citations by looking at their co-occurances, then we can group them into clusters to use at taxa for the extinction-innovation modelling.

Here's a table of pairwise counts for all citations in our sample:

```{r}
library(tidyverse)
library(tidytext)
# count how many times each pair of citations occurs together in a journal article.
# this is very slow, and produces a data frame with 27 million rows...
library(widyr)
title_word_pairs <-
  xx %>%
  pairwise_count(.out,
                 title,
                 sort = TRUE,
                 upper = FALSE)

title_word_pairs
```

With these pairwise counts we can draw a network plot to get a sense of how meaningful the co-occurances are:

```{r}

# network plot of citations: co-citations
library(ggplot2)
library(igraph)
library(ggraph)
library(tidygraph)

set.seed(1234)
title_word_pairs %>%
  filter(n >= 25) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "cyan4") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE,
                point.padding = unit(0.2, "lines")) +
  theme_void()
```

We can do community detection on this network to identify clusters of similar citations, according to the frequency of their co-occurance:

```{r}

title_word_pairs %>%
  filter(n >= 25) %>%
  graph_from_data_frame() %>%
  as_tbl_graph() %>%
  mutate(community = as.factor(group_edge_betweenness())) %>%
  ggraph(layout = 'kk') +
  geom_edge_link(aes(alpha = ..index..), show.legend = FALSE) +
  geom_node_point(aes(colour = community), size = 7) +
    geom_node_text(aes(label = name), repel = TRUE,
                point.padding = unit(0.2, "lines")) +
  theme_graph()

```

And here we can get the community membership in a data frame with the citations:

```{r}
citations_community <- 
  title_word_pairs %>% 
 filter(n >= 15) %>%
  graph_from_data_frame() %>%
  as_tbl_graph() %>%
  mutate(community = as.factor(group_edge_betweenness()))  %>% 
  as_tibble %>% 
  arrange(community)

# how many unique communities here?
length(unique(citations_community$community))
```

We need columns of the community number, the earliest publication year in the community and the most recent citation year in the community (lineage-style data). We could also give each community a broader discipline category. ). In this case, only having ten communities (taxa) might be too small, as we wouldn’t really be capturing the range of different communities.

```{r}
  # to occurance format
xx_with_communities_occ_format <- 
xx %>% 
  left_join(citations_community, 
            by = c(".out" = "name")) %>% 
  filter(year >= 1975) %>% 
  mutate(time_unit = year,
         Species = community) %>% 
  mutate(citing_paper = paste0(authors, 
                               " ",
                               year,
                               " ",
                               title,
                               " ",
                               journal)) %>% 
  select(citing_paper,
         .out, 
         time_unit, 
         Species) %>% 
  group_by(.out) %>%
  mutate(min_age = min(time_unit), 
         max_age = max(time_unit),
         Status = "extinct") %>% 
  distinct(.out, 
           min_age, 
           max_age, 
           Status,
           Species)

# drop items that are only cited in a single year
xx_with_communities_occ_format_multi_year <- 
  xx_with_communities_occ_format %>% 
  rename(Specimen = .out
         ) %>% 
    select(Specimen,
           Status,
           min_age,
           max_age,
           Species) %>% 
  mutate(span =  max_age - min_age) %>% 
  filter(span != 0) %>% 
  select(-span)

# convert to to lineage format
xx_with_communities_lineage_format_multi_year <- 
  xx_with_communities_occ_format_multi_year %>% 
  group_by(Specimen) %>% 
  summarise(first_year = min(min_age),
            last_year = max(max_age),
            Species = Species)

head(xx_with_communities_lineage_format_multi_year)
```

## Basic explorations of the citation clusters

We can look to see how dominant each citation community is over time to see trends in how archaeologists cite literature.

Here we see the number of citation communities in the literature over time:

```{r}
explore_citation_communities <- 
xx %>% 
  left_join(citations_community, 
            by = c(".out" = "name")) %>% 
  filter(year >= 1975) %>% 
  mutate(time_unit = year,
         Species = community) %>% 
  mutate(citing_paper = paste0(authors, 
                               " ",
                               year,
                               " ",
                               title,
                               " ",
                               journal)) %>% 
  select(citing_paper,
         .out, 
         time_unit, 
         Species) %>% 
  filter(!is.na(Species))

# each community as a proportion of all communities in that year.

communities_per_year <- 
explore_citation_communities %>% 
  group_by(Species, 
           time_unit) %>% 
  tally() %>% 
  ungroup() %>% 
  group_by(time_unit) %>% 
  mutate(perc = n / sum(n) * 100) %>% 
  ungroup()

communities_per_year %>% 
  select(time_unit, Species) %>% 
  group_by(time_unit) %>% 
  tally() %>% 
  ggplot(aes(time_unit,
             n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Number of citation communities per year")

```

Here we compute the percentage of each community in all of the cited works that we designate to a community, for each year. We only plot the communities that show a high range, because many show little variation over time.

We see some interesting trends:

- community 5, Shanks-Clarke-Binford-Hodder-Trigger (theory debates) dominate until about 2000. We also see a similar pattern with community 20, Semenov-Keely-Tringham (lithic usewear)   
- community 1, Bordieu-Giddens-Tilly-Gell-Latour-Ingold-Bender dominate after 2000, also dominant is community 2, Derrico-Eerkens-Henshilwood-Shennan (Pleistocene cultural evolution)    
- community 24, Stuiver on radiocarbon dating, appears in the mid-1990s, peaks in the mid-2000s, then drops steeply.    
- community 9, Bronk-Hogg-Reimer (radiocarbon calibrations) makes a low-profile appearance in the mid-1990s then breaks-out in the late 2000s

```{r}
# just plot the ones that have a wide range, who care about the flatliners
# get communities that range in percentage more than a certain amount 
communities_per_year_wide_range <- 
  communities_per_year %>% 
  group_by(Species) %>% 
  summarise(range = max(perc) - min(perc)) %>% 
  arrange(desc(range)) %>% 
  filter(range >= 10)

# plot communities over time. 
citation_communities_over_time <- 
communities_per_year %>% 
  filter(Species %in% communities_per_year_wide_range$Species) %>% 
  ggplot(aes(time_unit, 
           perc,
           colour = Species)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  scale_y_log10() +
  theme_minimal() +
  ggtitle("Change over time in citation communities in archaeological literature") +
  xlab("Year") +
  ylab("Percentage of all cited items in that year")

plotly::ggplotly(citation_communities_over_time)
```

We can make a histogram showing the distribution of the number of years each community was cited for (community lifespan):

```{r}
xx_with_communities_occ_format_multi_year_hist <- 
  xx_with_communities_occ_format_multi_year %>% 
    filter(!is.na(Species)) %>% 
  mutate(number_of_years_cited = max_age - min_age + 1)

# ggplot(refs_occurance_format_hist,
#          aes(number_of_years_cited)) +
#   geom_histogram()

hist(xx_with_communities_occ_format_multi_year_hist$number_of_years_cited,
     main = paste("Number of years each article in a citation community was cited for"),
     xlab = "Number of years cited")
```

## PyRate analysis

```{r}
library(pyrater)

# can't do it with occurance format: Error: cannot allocate vector of size 29.5 Mb 
# so let's use the lineage format, where each citation only appears once
refs_lineage_format <- 
  xx_with_communities_occ_format_multi_year %>% 
  filter(!is.na(Species)) %>% 
  group_by(Species) %>% 
  summarise(first_year = min(min_age),
            last_year = max(max_age)) %>% 
  mutate(taxa = "NA") %>% 
  select(taxa, 
         Species,
         first_year,
         last_year)

# must be  c("clade", "species", "min_age", "max_age")

refs_species_pyrate_data <- lineage_pyrate(refs_lineage_format)

refs_species_pyrate_data$species_pyrate_data$clade <- 0

#  We can preview each data frame:
lapply(refs_species_pyrate_data, head)
```

We write a text file to be used with the PyRate analysis later on. This file will appear in our current working directory

```{r, eval=TRUE}
write.table(refs_species_pyrate_data$species_pyrate_data,
            "lineage_pyrate_data.txt",
            quote=FALSE,
            sep="\t",
            row.names = FALSE)
```

### Summary Plots

Prior to starting the PyRate analysis, it is good practice to get a sense of the data by developing a series of plots.  Using the function below, the following plots are created

1. Log Lineages Through Time
2. Cumulative Diversity Through Time
3. Average Lifespan Through Time
4. Histogram of Lifespans

#### Citation Data (Lineage Data)

```{r, eval=TRUE, fig.height=10, fig.width=14}
# The numeral indicates the x-axis increments with 
# the number 5 indicating 5-year time bins
# windows() # if interactive
summary_plots(refs_species_pyrate_data$species_pyrate_data, 5) 
```

### Choosing the best underlying model

From https://www.nature.com/articles/palcomms201619

```{r, eval=FALSE}

# set the number of generations. 10,000,000 generations = 10 hrs

n_generations <- 1000000

library(pyrater)

# birth-death model
system(
    paste0(
      "C:/python27/python.exe",
      " ",
      system.file("Pyrate/PyRate.py", package = "pyrater"),
      " -d ",
      " lineage_pyrate_data.txt ",
      " -A 2 ",
      " lineage ",
      " -mBDI ",
      0,
      " -n ",
      10000
    ),
    intern = TRUE, wait = FALSE
  )
# elapsed time:   sec

# immigration-death model
system(
    paste0(
      "C:/python27/python.exe",
      " ",
      system.file("Pyrate/PyRate.py", package = "pyrater"),
      " -d ",
      " lineage_pyrate_data.txt ",
      " -A 2 ",
      " lineage ",
      " -mBDI ",
      1,
      " -n ",
      10000
    ),
    intern = TRUE, wait = FALSE
  )
#elapsed time:  " 
```

This analysis uses the default settings of samples every 1,000 generations for 10,000,000 generations.  Details about changes to PyRate default settings can be found in Silvestro et al. 2014. If access to a computing cluster is available, this is the preferred option. To reduce the run time of the analysis, simply adjust the number of generations from 10 million to 1 million.

# Visualizing the Rates of Origination and Extinction

Once the BDMCMC runs are finished, one can easy view the a rate through time plot by using the following commands:

```{r, eval=FALSE}

pyrate_plot(path_to_python = "C:/python27/python.exe",
            input_file = "pyrate_mcmc_logs/lineage_pyrate_data_0_BD_marginal_rates.log")
```

We can redraw these plots in R:

```{r, eval=TRUE}
source(paste0(here::here(), "/pyrate_mcmc_logs/lineage_pyrate_data_0_BD_marginal_rates_RTT.r"))

age_rev <-
  2017 + age # adjust for the most recent date in each of the datasets

# Origination
plot(
  age_rev,
  L_hpd_M95,
  type = 'n',
  ylim = c(0, max(L_hpd_M95)),
  xlim = c(min(age_rev), max(age_rev)),
  xlab = "Years",
  ylab = "Rate of Event Per Lineage Per Time Unit",
  main = "Origination"
)
plot_RTT(age_rev, L_hpd_M95, L_hpd_m95, L_mean, "navyblue")

# Extinction
plot(
  age_rev,
  M_hpd_M95,
  type = 'n',
  ylim = c(0, max(M_hpd_M95)),
  xlim = c(min(age_rev), max(age_rev)),
  xlab = "Years",
  ylab = "Rate of Event Per Lineage Per Time Unit",
  main = "Extinction"
)
plot_RTT(age_rev, M_hpd_M95, M_hpd_m95, M_mean, "red")

# Net Diversification
plot(
  age_rev,
  R_hpd_M95,
  type = 'n',
  ylim = c(min(R_hpd_m95), max(R_hpd_M95)),
  xlim = c(min(age_rev), max(age_rev)),
  xlab = "Years",
  ylab = "Rate of Event Per Mineage Per Time Unit",
  main = "Net Diversification"
)
plot_RTT(age_rev, R_hpd_M95, R_hpd_m95, R_mean, "darkgreen")
```